---
title: "NYPD_Shooting_Case_Study"
author: "Steven Sharp"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Science as a Field NYPD shooting case study project



```{r Import_libs}
library(tidyverse)
library(lubridate)
library(usmap) #import the package
library(ggplot2) #use ggplot2 to add layer for visualization
library(leaflet)

```


## Data Sources 
### Importing data 
```{r GetData}
url_in <- str_c("https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD")
NYPD_shootings <- read.csv(url_in[1])

```


 

## Tidy up the data
### remove unwanted data

I what to remove data older than 10 years. I want to look at location and time of day as factors to investigate. I will group locations by Borough and times. I will do further analysis to see if their are trends in the data


```{r tidy}


# Remove missing values from NYPD_shootings
NYPD_shootings <- na.omit(NYPD_shootings)

#change date object
NYPD_shootings$OCCUR_DATE <- as.Date(NYPD_shootings$OCCUR_DATE, format = "%m/%d/%Y")

#look at the last 10 years
NYPD_shootings <- NYPD_shootings %>% filter(OCCUR_DATE > ymd("2012-12-31"))


```



review data 
```{r inspect}

head(NYPD_shootings,5)
```

### Group data 
I want to see if there are trends in the locations or times (or both). So make two variables for day and night and filter those records with those attributes. I define day from 6am-6pm everything else as night and 

```{r filter_group}


NYPD_shootings_day <- NYPD_shootings %>% filter(OCCUR_TIME  >= "06:00:00" & OCCUR_TIME <= "17:59:59")

NYPD_shootings_night <- NYPD_shootings %>% filter(OCCUR_TIME >= "18:00:00" & OCCUR_TIME <= "5:59:59")


```

## Show two different visualization
inspecting data for trend or insight

```{r graph data}



library(leaflet)

# Set the bounding box coordinates for New York City
min_lat <- 40.5774
max_lat <- 40.9176
min_long <- -74.15
max_long <- -73.7004

# Create a leaflet map object centered around New York City
map <- leaflet() %>%
  setView(lng = mean(c(min_long, max_long)), lat = mean(c(min_lat, max_lat)), zoom = 11) %>%
  addTiles()

# Add markers for nighttime shootings
map_night <- addCircleMarkers(map, data = NYPD_shootings_night, lng = ~Longitude, lat = ~Latitude,
                              color = "black", radius = 0.05, fillOpacity = 0.01)

# Add markers for daytime shootings
map_day <- addCircleMarkers(map_night, data = NYPD_shootings_day, lng = ~Longitude, lat = ~Latitude,
                            color = "blue", radius = 0.05, fillOpacity = 0.01)

# Print the map
map_day


```


From the above map it looks like I should group in Boroughs, and that night time shootings are more frequent.


## Step 2
### Ploting to compare day time and night time shootings
```{r graph data2}

# Combine the "Day" and "Night" datasets into a single dataframe
stacked_data <- rbind(transform(NYPD_shootings_day, Period = "Day"), transform(NYPD_shootings_night, Period = "Night"))

#  code for the stacked bar chart
ggplot(data = stacked_data, aes(x = fct_rev(fct_infreq(BORO)), fill = Period)) +
  geom_bar(color = "blue", position = "stack") +
  ggtitle("Number of shootings in different BOROUGH") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Borough") + ylab("Shooting Count") +
  coord_flip() +
  theme(legend.position = "top")


```
From the chart above it does look like night time shootings are more prevalent

## Step 3 Model 
lets looking to see if there is a trend to the data 

```{r model}





# Calculate the number of nighttime shootings by date
nighttime_counts <- aggregate(OCCUR_TIME ~ OCCUR_DATE, data = NYPD_shootings_night, FUN = length)

# Create a model using linear regression
lm_model <- lm(OCCUR_TIME ~ OCCUR_DATE, data = nighttime_counts)
trend_line <- data.frame(OCCUR_DATE = nighttime_counts$OCCUR_DATE,
                         OCCUR_TIME = predict(lm_model))

# Plot nighttime shootings with dates and trend line
ggplot(nighttime_counts, aes(x = OCCUR_DATE, y = OCCUR_TIME)) +
  geom_point(color = "black") +
  geom_line(data = trend_line, aes(x = OCCUR_DATE, y = OCCUR_TIME), color = "red") +
  scale_x_date(date_labels = " %Y", date_breaks = "1 year") +
  labs(x = "Date", y = "Number of Night time Shootings",
       title = "Nighttime Shootings over Time with Trend Line") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

summary(lm_model)

# Calculate the number of daytime murders by date
daytime_counts <- aggregate(OCCUR_TIME ~ OCCUR_DATE, data = NYPD_shootings_day, FUN = length)

# Create a model using linear regression
lm_model2 <- lm(OCCUR_TIME ~ OCCUR_DATE, data = daytime_counts)
trend_line2 <- data.frame(OCCUR_DATE = nighttime_counts$OCCUR_DATE,
                         OCCUR_TIME = predict(lm_model))

# Plot daytime shootings with dates and trend line
ggplot(daytime_counts, aes(x = OCCUR_DATE, y = OCCUR_TIME)) +
  geom_point(color = "blue") +
  geom_line(data = trend_line2, aes(x = OCCUR_DATE, y = OCCUR_TIME), color = "red") +
  scale_x_date(date_labels = "%b %Y", date_breaks = "12 month") +
  labs(x = "Date", y = "Number of Day time Shootings",
       title = "Daytime Shootingss over Time with Trend Line") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

summary(lm_model2)

```
```{r analisys}
# Line 1 parameters
slope1 <- 2.183e-04
intercept1 <- -1.494

# Line 2 parameters
slope2 <- 1.403e-04
intercept2 <- -5.745e-01

# Calculate the x-coordinate of the intercept
intercept_x <- (intercept2 - intercept1) / (slope1 - slope2)

# Calculate the y-coordinate of the intercept
intercept_y <- slope1 * intercept_x + intercept1

# Create a plot
plot(x = c(11700, 11810), y = c(-10, 10), type = "n", xlab = "X", ylab = "Y")

# Plot line 1
abline(a = intercept1, b = slope1, col = "red", lwd = 2)

# Plot line 2
abline(a = intercept2, b = slope2, col = "blue", lwd = 2)

# Mark the intercept point
points(intercept_x, intercept_y, col = "green", pch = 16, cex = 2)

# Add a legend
legend("topleft", legend = c("Line 1", "Line 2", "Intercept"), col = c("red", "blue", "green"), lty = 1, lwd = 2, pch = c(NA, NA, 16), cex = 0.8)

```



From the above trend lines and the summary, both day and night for the past ten years are trending **up**, with the day increasing at a faster rate than nigh time Shootings.

## Step 4 
## Addressing Bias

Bias that might affect my approach:
- There might be Bias in the data, how it's collected, what constatues a shooting (do you need multiple witnesses? )
- I chose time and Boroughs as variables because I was interest to see if time of day and grouping had any correlation on occurrence. 
- I'm not from a large city so I may not appreciate how many people may be working and traveling at night. 



