---
title: "UseRmd2CreateDoc"
author: "Steven Sharp"
date: "2023-05-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

### A Bulleted List

* Item A
* Item B
  + Item B1
  + Item B2
  
### Numbered List
1. Item 1
2. Item 2
3. Item 3
    + Item 3a
    + Item 3b
    
### Bold and italics

*italic*   **bold**

_italic_   __bold__

# Header 1
## Header 2
### Header 3
#### Header 4

### R code without showing code
```{r cars, echo=FALSE}
summary(cars)
```

### R code showing 
```{r cars2, echo=TRUE}
summary(cars)
```


## TABLE
table with 3 columns

| Left-aligned | Center-aligned | Right-aligned |
| :---         |     :---:      |          ---: |
| First row    | R1 C2          | R1 C3         |
| Second row   | R2 C2          | R2 C3         |
| Third row    | R3 C2          | R3 C3         |

### Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

### Session Info

```{r sessionInfo, echo=TRUE}
utils::sessionInfo()
```
